---
title: "Getting Started with Stan"
author: "Gabriel J. Odom, PhD, ThD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Overview
So you want to do some Bayesian inference? Well, you'll need some software. If you are reading this, you should already have [R](https://cran.r-project.org/) and [RStudio](https://www.rstudio.com/products/rstudio/download/) installed on your computer (**MAKE SURE YOU HAVE `R` VERSION 4.0 OR HIGHER**). This document will help you get started with a new piece of statistical software: [Stan](https://mc-stan.org/). Here are the steps to install this new software.



# Step 1: Configure your Computer for C++
Unlike R, code in Stan requires that your computer be able to compile code.


## Compiling Code on Windows
These steps are from [here](https://github.com/stan-dev/rstan/wiki/Configuring-C---Toolchain-for-Windows). You first want to ensure that you have the right Windows environment variables for R:
```{r}
Sys.getenv("BINPREF")
```

The code above should return `""` at this point in the install process.

Now, install [Rtools](https://cran.r-project.org/bin/windows/Rtools/). You will also set another R environment variable using the code below:
```{r}
writeLines('PATH="${RTOOLS40_HOME}\\usr\\bin;${PATH}"', con = "~/.Renviron")
```

After you run this code, restart RStudio, and check that your system path has been updated.
```{r, eval=TRUE}
Sys.which("make")
```

The code above should return `"C:\\rtools40\\usr\\bin\\make.exe"`.


## Compiling Code on Mac
These steps are from [here](https://github.com/rmacoslib/r-macos-rtools). Because this process downloads Xcode, it may take some time (and I recommend that you use an ethernet internet connection if possible).

